Optimal Instantiation of Abstract Workflows using Logic and Circuits

### 2009-12-22 publicaciones/eswc10/paperESWC10.tex
paper submiteado

  \documentclass{llncs}
\usepackage{amsmath}
\usepackage{txfonts}
\usepackage[boxed]{algorithm2e}
\usepackage{amssymb}
\usepackage[pdftex]{graphicx}
\usepackage{enumerate}
\DeclareGraphicsExtensions{.jpg, .png, .mps, .bmp, .pdf}
\usepackage{epsfig}
\usepackage{subfigure}
\newtheorem{Definition}{Definition} 
\newtheorem{Example}{Example} 
\newtheorem{Theorem}{Theorem} 
\newtheorem{Proof}{Proof} 
\newcommand{\denselist}{\topsep 0pt \itemsep -4pt}
\newcommand{\tup}[1]{\langle #1 \rangle}
\newcommand{\vvec}[1]{\mathbf{#1}}
\newcommand{\join}{\bowtie}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\body}{{body}}
\newcommand{\head}{{head}}
\newcommand{\qrule}{:\!\!-}
\newcommand{\arc}{\text{arc}}
\newcommand{\Theory}[1]{T(#1)}
\newcommand{\mcdsat}{\textsc{McdSat}}
\newcommand{\minicon}{{MiniCon}}
\newcommand{\Omit}[1]{}
\newcommand{\citeX}[1]{\citeauthor{#1}~\citeyear{#1}}

\begin{document}
\allowdisplaybreaks
\title{Selecting the Maximum-Utility  Web Services  Using Circuits}
\author{Daniel Izquierdo \inst{1}  and Blai Bonet \inst{1}  and Mar\'{\i}a-Esther Vidal \inst{1}}

\institute{Universidad Sim\'on Bol\'{\i}var\\
                Caracas, Venezuela\\
                \{idaniel,bonet,mvidal\}ldc.usb.ve} \maketitle
\begin{abstract}
In Service Oriented Architectures (SOA), concrete services are described in terms of functionality   and QoS parameters, while software applications are specified as workflows of abstract services and non-functional criteria that establish the  application desirable functionality and  expected quality, respectively.  During workflow execution time,  abstract services  are rewritten 
with the concrete services  that meet the workflow  functional and non-functional restrictions. Since, service selection is done on-the-flight and the space of concrete services may be very large, efficiency and scalability are required.  In this paper we propose the {\it McdSat}$^c$ framework to efficiently solve the problem of service selection in presence of a large number of concrete services.
First, we use the Local As View approach to describe the functionality  of the concrete services in terms of views of abstract services, the quality as an overall utility function that combines the different QoS values that describe the service, and  workflows as  conjunctive queries on abstract services. Based on this representation, we 
model the problem of service selection as a known problem in the area of integration systems referred as query rewriting using views.  We encode the service selection problem as a propositional theory whose models correspond to the composition of the services that implement the abstract workflow and have the same cost as the corresponding composition; known relationships between d-DNNF theories and arithmetic circuits are exploited to provide an efficient and scalable solution.  We report on the performance of our proposed approach, and we observe that  {\it McdSat}$^c$  efficiently scales up to large instances of the service selection problem.

\end{abstract}                
\section{Introduction}
Under the umbrella of the Semantic Web and supported by Service Oriented architectures, the number of Web data sources and services has exploded during the last few years.  For example,  currently, the molecular biology databases
collection includes 1,170 databases~\cite{Galperin09}, that is
95 more than the previous year~\cite{Galperin2008} and 110 more than two years ago~\cite{Galperin2007};  tools
and services as well as the number of instances published by these resources, follow a similar progression~\cite{Benson07}. Thanks
to this wealth, users rely more on various
digital tasks such as data retrieval from public data sources and
data analysis with Web tools or services organized in
complex workflows.

In order to design an application, users usually specify the functionality in terms of a workflow of abstract services, while the desirable quality  is defined by restricting the values of the QoS parameters that describe the concrete available services. To evaluate the proposed workflow,  abstract services are instantiated on concrete services in a way restrictions on the QoS parameters are satisfied.  However, since the number of available sources has an impact on the performance of this task, techniques that enable us to efficiently scale up to a large number of services are required.

In this paper we aim to provide a solution to this problem and provide techniques that efficiently traverse the space of available services and select combinations that meet the functional and non-functional requirements that describe a designed application.  We name our proposed solution {\it MCDSat}$^c$.

In {\it MCDSat}$^c$, concrete service functionalities are described as views on abstract services and workflows are represented as conjunctive queries on abstract services; this service representation is similar to the one generated in a semi-automatic way for the {\cal DEIMOS} system ~\cite{AmbiteISWC09}.  In addition, rules that define workflows and  concrete services in the {\it MCDSat}$^c$ catalog,  are created in a way that input and output restrictions of the combined abstract services are satisfied. To represent the QoS measures, each concrete service description is annotated with a real number that represents an overall value of the QoS measures of this service.  Finally, {\it MCDSat}$^c$ models the web service selection problem as the problem of query rewriting using views \cite{halevy:survey}. This problem is important in the context of data integration
\cite{Chen05,JaudoinPRST05}, and query optimization and data maintenance \cite{AfratiLU07,levy:bucket}, and several approaches have been defined to efficiently scale up to a large number of views~\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}. In this paper we propose the {\it MCDSat}$^c$ framework which extends the approach proposed in ~\cite{arvelo:aaai06},  with the capability of dealing with constants and identifying the composition of services that maximizes a given utility function.

{\it MCDSat}$^c$  translates an instance of the service selection problem into the problem of enumerating the models of a propositional theory that encodes information about the workflow, the concrete services and the conditions that need to be satisfied in order to generate a composition that maximizes the utility function. This theory is represented as a CNF formula which is compiled into a deterministic
and decomposable negation normal form d-DNNF representation, from which the models that maximizes the overall utility function are obtained in linear time using modern SAT techniques~\cite{darwiche:dnnf}.   
We empirically studied the quality of the {\it MCDSat}$^c$ on variety of benchmarks and we observed  that the proposed technique provides an efficient solution to the service selection problem which is able to scale up to large workflows as well to a large number of views.

 This paper is comprised of six additional sections. The next section motivates our approach by using an example, and  we summarize existing related work in section 3. Section 4 describes the {\it MCDSat}$^c$ system architecture, and  the experimental study is reported in section 5. 
 In section 6 we present the formalization of the web service selection problem as a Propositional Logic theory, and finally, section 7 outlines our conclusions and future work.

\section{Motivating Example}
Consider the abstract and concrete services in Table ~\ref{table:services}  which represent applications in the flight domain.

\begin{table}
\begin{tabular}{| l| l |}
\hline
Abstract Service & Description \\
\hline
$fight(x1,x2)$ & relates cities x1 and x2 if there is a flight between them\\
\hline
$uscity(x1)$ &  determines if a city x1 is a US city\\
\hline \hline
Concrete  Service & Description \\
\hline
$national(x1,y1)$&  returns  two US cities that are connected by  direct flights \\
\hline
$oneway(x1,y1)$ & returns two cites that are connected by one way direct flights \\
\hline
$onestop(x1,y1)$ &  returns  two cites that are connected by one-stop flights \\
\hline
$ flightToCCS(x1,"CCS")$ & returns a city that is connected to  Caracas by a direct flight\\
 \hline
$ onestopToCCS(x1,y1)$ & returns two cites x1 and y1, such that there is  one-stop flight\\ 
 & from city x1 to Caracas stopping at city y1.\\
 \hline
 $fromWAS("WAS",y1)$ &  returns a city y1 to which flights from Washington arrive. \\
 \hline
\end{tabular}
\label{table:services}
\caption{Abstract and Concrete Services}
\end{table}

Using the  Local As View (LAV) approach~\cite{AmbiteISWC09}, concrete services in Table 1 are semantically described as views on the abstract services {\it flight} and {\it uscity} as follows:

\begin{alignat*}{1}
national(x1,y1)\ &\qrule\ flight(x1,y1),\,  uscity(x1),\,  uscity(y1)\,. \\
oneway(x2,y2)\ &\qrule\ flight(x2,y2)\,. \\
onestop(x3,z3)\ &\qrule\ flight(x3,y3),\, flight(y3,z3)\,. \\
flightToCCS(x2,"CCS")\ &\qrule\ flight(x2,"CCS")\,. \\
onestopToCCS(x3,y3)\ &\qrule\ flight(x3,y3),\, flight(y3,"CCS")\,. \\
fromWAS("WAS",y3)\ &\qrule\ flight("WAS",x3)\,. \\
\end{alignat*}

Now suppose a user is interested in implementing a workflow able to retrieve the one-stop round trip flights from US cities to any city in the world, such that flights can stop at any city. 
Consider the following conjunctive query that represents the workflow that defines this request in terms of abstract services. This workflow was defined in way the input parameters of each abstract service are bound by  constants in the workflow, or by attributes projected out by  previous abstract services.

\begin{alignat*}{1}
w(x,y1,y2,y3)\ &\qrule\ flight(x,y1),\, flight(y1,y2),\,  flight(y2,y3),\,  flight(y3,x),\,  \\
                   \ & uscity(x)\,. \\
\end{alignat*}

Implementations of this workflow correspond to compositions of the concrete services where each concrete service can implement a subset of abstract services of the workflow, but each abstract service can be implemented by exactly one concrete service. The following composition of concrete services corresponds to one of the XX implementations of the workflow:

\begin{alignat*}{1}
wi1(x,y1,"CCS",y3)\ &\qrule\ national(x,y1),\, flightToCCS(y1,"CCS"),\,  oneway("CCS",y3),\,  \\
\ & national(y3,x)\,.  \\
\end{alignat*}

However, note that the following compositions are not valid:

\begin{alignat*}{1}
wi2(x,y1,"CCS"y3)\ &\qrule\ national(x,y1),\, flightToCCS(y1,"CCS"),\, \\
\ & fromWAS("WAS",y3),\,  national(y3,x)\,. \\
wi3(x,y1,y2,y3)\ &\qrule\ national(x1,y1),\, onestop(x1,y2),\,  oneway(y2,y3),\,  \\
\ & national(y3,x)\,. \\
\end{alignat*}

The composition {\it wi2} is not valid because  it maps the workflow variable
{\it y2} into constants {\it "CCS"} and {\it "WAS"} which denote different cities.
On the other hand, the composition {\it wi3} does not implement the workflow because the concrete service  {\it onestop} does not receive as input or produce as output the city where the flight stops (represented by the variable {\it y1}); thus, it is not possible to ensure that the city where this flight stops and the end city  retrieved by the service {\it national}, are the same.

To avoid producing invalid workflow implementations, it is required to handle constants in a way that different constants are not mapped into
each other either directly or indirectly (via transitivity). In addition,  all the attributes in the workflow output  or needed to join other services in the workflow, have to be produced by the selected concrete service.

In addition, suppose each concrete service is annotated with a utility function that aggregates the values of different QoS that characterize the behavior of the service. Then, since the space of valid service compositions of an abstract workflow can be very large, we require a technique able to identify the ones that maximize the utility function  without having to enumerate the whole set of valid compositions. In this paper, we propose a Propositional logic-based  approach that takes advantage of the  power of  modern SAT solvers, to efficiently enumerate the service compositions that correspond to valid implementations of a given abstract workflow, as well as the compositions that maximize a given utility function.

\section{Related Work}
In this section we summarize existing approaches that provide solutions to the problems of service selection, knowledge representation and query rewriting.

\begin{description}
\item[Service Selection Solutions] \mbox{}\\
The problem of selecting the services that implement an abstract workflow and best meet the QoS-based criteria  is known as the QoS-aware service selection or composition problem, which has been shown to be NP-hard~\cite{Hiroshi2008}. This problem is a combinatorial optimization problem and several heuristics have been proposed to find a relatively good solution in a reasonably short period of time.  A  distance metric-based heuristic  to drive a backward search algorithm  is proposed~\cite{rahmani08}; this metric induces an order of the services in a way that sink nodes are unlikely to be visited. In ~\cite{berardi05,berardi08,berardi06}, services and workflows are described in terms of deterministic finite state machines that are encoded as a Description Logic theory whose models correspond to solutions of the problem;  although reasoning methods for Description Logics formalisms could be exploited, scalability or performance of the proposed solution has not been reported.  ~\cite{myoung08} propose a constraint-based approach that encodes the non-functional permissible values as a set of constraints whose violation needs to be minimized; to traverse the space of possibly optimal solutions, a hybrid algorithm that combines the tabu search and simulated annealing meta-heuristics is implemented; experimental results show that the proposed solution is able to scale up to a large number of services and abstract processes.  In ~\cite{cardellini07} the QoS-aware service composition problem is encoded as a Linear Programming problem providing a scalable solution to the problem.  In ~\cite{Hiroshi2008}  this problem  is defined as a multi-objective optimization problem where the different QoS parameters are considered equally important and there is not an aggregated function to combine all of them; a genetic-based algorithm is proposed to identify a set of non-dominated service compositions that best meet all the QoS parameters.  ~\cite{alrifaiR09} propose a two-fold solution that uses a hybrid integer programming algorithm to find the decomposition of global QoS into local constraint, and then,  selects the services that best meet the local constraints.   
Recently,  two new planning-based approaches have been proposed~\cite{kuterG09,sohrabiM09}.  ~\cite{kuterG09} extend the SHOP2 planning algorithm to select the trustworthy composition of services that implement a given OWL-S process model, while ~\cite{sohrabiM09}  propose a HTN planning-based solution where user preference metrics and domain regulations are used to guide the planner into the space of relevant compositions. Finally, ~\cite{lecue09} proposes a genetic-based algorithm to identify the composition of services that best meet the quality criteria for a set of QoS parameters.

Although these solutions are able to efficiently solve the optimization problem and scale up to a large number of abstract processes, none of them are tailored to semantically describe services in terms of the abstract process, or use these descriptions to identify the services that  implement a given workflow and best meet user non-functional criteria.

\item[Knowledge Compilation Languages] \mbox{}\\
Knowledge compilation is the area in AI concerned with the
problem of mapping logical theories into suitable fragments
that make certain desired operations tractable
\cite{cadoli:compilation}. Different compilation languages have been defined, for instance,  
Ordered Binary Decision Diagrams (OBDDs)~\cite{bryant:obdd},  Negation Normal Form (NNF) \cite{barwise:handbook}, and Decomposable Negation Normal Form (DNNF)  \cite{darwiche:dnnf}.
In this work we make use of the properties of the deterministic DNNFs  (d-DNNF) \cite{darwiche:d-dnnfs} to provide an scalable and efficient solution to the service selection problem.

A Negation Normal Form (NNF) theory is  constructed from literals using only conjunctions
and disjunctions \cite{barwise:handbook}, and it can be represented as a directed acyclic graph
in which the leaves are labeled with literals and the
internal nodes are labeled with $\land$ and $\lor$;
see Fig.~\ref{fig:dnnf} for an example. An NNF is said to be decomposable (DNNF) \cite{darwiche:dnnf}
if for each conjunction $\bigwedge_i\phi_i$, the set of
variables in each conjunct are pairwise disjoint;
i.e,. $Vars(\phi_i)\cap Vars(\phi_j)=\empty$ for $i<j$.
A DNNF supports a number of operations in
polynomial time in the size of its DAG. For example,
we can test whether a DNNF is satisfiable by a single
bottom-up pass over its DAG in linear time. A DNNF is said to be deterministic (d-DNNF) \cite{darwiche:d-dnnfs} if for each disjunction $\bigvee_i\phi_i$, the disjuncts
are pairwise logically contradictory; i.e.,
$\phi_i\Rightarrow\neg\phi_j$ for $i<j$.
The NNF in Fig.~\ref{fig:dnnf}, for example, is decomposable
and deterministic. A d-DNNF supports the model counting and enumeration in polynomial time
in the size of its DAG.  {\it MCDSat}$^c$  exploits the properties of a d-DNNF theory, to efficiently provide a solution to the problem of enumerating the service compositions that correspond to models of the  theory, i.e., the service compositions that implement the abstract  workflow and best meet the utility function. DECIR ALGO SOBRE EL ALGORITMO QUE PRODUCE LOS MEJORES.

\begin{figure}
\centering
\includegraphics[width=8cm]{odd}
\caption{A decomposable and deterministic NNF.}
\label{fig:dnnf}
\end{figure}

\item[Query Rewriting Solutions] \mbox{}\\
A number of algorithms have been developed to find the rewritings of a given query; the most prominent being the bucket algorithm~\cite{levy:bucket},  the inverse rules algorithm ~\cite{duschka:answer,Qian96}, the minicon algorithm ~\cite{pottinger:minicon}, and the {\it MCDSat} ~\cite{arvelo:aaai06}. Generally, query rewriting algorithms work in two phases: first, they identify the views that rewrite at least one  subgoal of the query; second they combine the selected views to produce a rewriting. The main difference between existing approaches, is the criteria used to choose the relevant views and reduce the space of non-useful rewritings.

The bucket algorithm reduces the number of possibilities just considering each subgoal in the query in isolation, and selecting the views that are able to  produce at least the attributes projected by the query. Since, attributes involved in query joins are not verified, a large number of rewritings comprised of  Cartesian products may  be generated. 	
The Inverse Rules algorithm constructs a set of rules that invert the view definition and establish how to compute tuples for the database relations from tuples of the views. Similarly, to the bucket algorith it can produce a large number non-useful of rewritings.

The MiniCon algorithm  overcomes the limitations of the previous algorithms by identifying only views that rewrite a set of the query goals and that  can be combined with the rest of the query subgoals. The key idea is to identify the mappings between  the variables in each subgoal of the query to the variables  in one or more subgoals of the views, in a way that, join variables in the query are mapped to join variables in the body of a view or to the distinguished variables of the view. Mappings between variables and subgoals are represented in MiniCon Descriptions (MCD's)\cite{pottinger:minicon}.

>>>>

\documentclass{llncs}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{enumerate}
\usepackage{url}

\newcommand{\tup}[1]{\langle #1 \rangle}
\newcommand{\vvec}[1]{\mathbf{#1}}
\newcommand{\join}{\bowtie}
\newcommand{\R}{\mathcal{R}}
\newcommand{\Q}{\mathcal{Q}}
\newcommand{\qrule}{:\!\!-}
\newcommand{\mcdsat}{\textsc{McdSat}}
\newcommand{\minicon}{{MiniCon}}
\newcommand{\Omit}[1]{}

% flights
\newcommand{\flight}{\text{\it flight}}
\newcommand{\UScity}{\text{\it uscity}}
\newcommand{\national}{\text{\it national}}
\newcommand{\oneway}{\text{\it one-way}}
\newcommand{\onestop}{\text{\it one-stop}}
\newcommand{\flightPA}{\text{\it flight-to-pa}}
\newcommand{\onestopPA}{\text{\it onestop-to-pa}}
\newcommand{\fromNY}{\text{\it from-ny}}
\newcommand{\PA}{\text{PA}}
\newcommand{\NY}{\text{NY}}
\newcommand{\AL}{\text{AL}}

\begin{document}

\allowdisplaybreaks
\title{Optimal Instantiation of Abstract Workflows using Logic and Circuits}
\author{Daniel Izquierdo \and Mar\'{\i}a-Esther Vidal \and Blai Bonet}
\institute{Departamento de Computaci\'on \\
           Universidad Sim\'on Bol\'{\i}var \\
           Caracas 89000, Venezuela \\
           \url{{idaniel,bonet,mvidal}@ldc.usb.ve}}
\maketitle

\begin{abstract}
The typical Service Oriented Architecture consists of two main layers.
The concrete layer with concrete services whose functionalities are
described in terms of pre- and post-conditions and non-functional
properties in terms of QoS parameters, and the abstract layer with
software applications whose functionalities are described in terms
of abstract workflows and non-functional properties in terms of
restrictions on QoS.
During the execution of a workflow, the abstract services are instantiated
into concrete services that meet the functional and non-functional
requirements. This instantiation, that must be done on-the-flight,
consists in a search on a combinatorial space of possibilities.
In this paper, we propose a framework to efficiently solve the
instantiation problem that is firmly grounded on logic.
The framework adopts the Local-As-View approach in which
the functionality of concrete services is described using views
of abstract services, the quality of an instantiation as an
overall utility function that combines the different QoS parameters,
and the abstract workflows as conjunctive queries on abstract services.
Using this representation, the problem of workflow instantiation 
is cast as a problem of query rewriting, from the area of integration
systems.
Then, building on related work, we devise an encoding of the
workflow instantiation problem as a logical theory whose models
are in correspondence with the instantiations of the workflow,
and the best ranked models are in correspondence with the
optimal instantiations of the workflow.
Thus, by exploiting known properties of logical theories in 
d-DNNF format, we provide an efficient and scalable solution
to the workflow instantiation problem.
The approach does not only scale up to large instances, as the
experimental results show, but is also sound and complete since,
being founded on logic, is amenable to formal analysis.
\end{abstract}

\section{Introduction}

Under the umbrella of the Semantic Web, and supported by Service Oriented
Architectures (SOA), the number of Web data sources and services has exploded
during the last few years. For example, the molecular biology database
collection currently contains 1,170 databases~\cite{Galperin09}, a number
that is 95 more than the previous year~\cite{Galperin2008} and 110 more
than two years ago~\cite{Galperin2007}; tools and services, as well as the
number of instances published by these resources, follow a similar
progression~\cite{Benson07}. Thanks to this wealth of resources, users'
tendency is to rely more on automatic methods for handling them such as
data retrieval from public sources and analysis using Web
tools or services composed in complex workflows.

The typical SOA consists of two layers. The concrete layers that is
made of concrete services whose functionalities are described in
terms of pre- and post-conditions and non-functional properties in
terms of QoS parameters, and the abstract layer that is made of 
software applications whose functionalities are described in terms
of abstract workflows and non-functional properties in terms of
restrictions on QoS.
The execution of an abstract workflow involves the instantiation
of the abstract services into concrete services that meet the
functional and non-functional requirements. This instantiation
process can be seen as a search for a target instantiation on
the combinatorial space of all valid instantiations.
Thus, one is interested in efficient techniques for performing
this search that are able to scale up as the number of concrete
services or the complexity of the workflow increases.
We call the problem of instantiating a given abstract workflow
with concrete services, from a given pool of concrete services,
such that certain QoS demands are meet, the Workflow Instantiation
Problem (WIP).

In this paper, we consider a restricted version of WIP that adopts
the Local-As-View (LAV) approach \cite{levy:bucket}.
In LAV, all elements in a problem are specified with a common language
that is grounded on \emph{abstract services} such that concrete
services are described as views of abstract services, the quality
of an instantiation as an overall utility function that combine the
different QoS parameters, and the abstract workflow as a conjunctive
query on the abstract services; this representation is similar to the
one that is semi-automatically generated for the DEIMOS system
\cite{AmbiteISWC09}.
In this version of WIP, the rules that define workflows (conjunctive queries)
and concrete services (views) are created in a way that all the 
functional restrictions on pre- and post-conditions of services
and their combinations are satisfied, and the QoS measures are
represented by annotating each concrete service description with
a real number that represent the overall QoS utility of the service.

Under these assumptions, the WIP can be cast as the well-known Query
Rewriting Problem (QRP) for LAV which is central to integration systems
\cite{halevy:survey}. The QRP consists of a conjunctive query that must be
answered in terms of views in which the query and the views are described
using LAV with abstract relations.
This problem is important in the context of data integration
\cite{Chen05,JaudoinPRST05}, and query optimization and data maintenance
\cite{AfratiLU07,levy:bucket}, and several approaches have been defined
that scale to a large number of views
\cite{arvelo:aaai06,pods:DuschkaG97,sac:DuschkaG97,levy:bucket,pottinger:minicon}.

The recent approach of Arvelo et al.\ \cite{arvelo:aaai06} is based 
on the efficient enumeration of models of a propositional logic theory. 
Given a QRP, a logical theory is constructed such that each model of
the theory encodes a valid rewriting, and thus all valid rewritings
are obtained by enumerating all models of the theory. This enumeration
can be efficiently performed if the logical theory is in certain 
normal form called deterministic and decomposable negation normal
form (d-DNNF) \cite{darwiche:d-dnnfs}. Thus, the approach consists in
transforming (called compiling in the field of knowledge compilation)
the logical theory into d-DNNF format for enumerating its models efficiently.

Yet d-DNNF theories not only support the efficient enumeration of
models but other operations too. Among them, the enumeration of
the best ranked models can also be performed efficiently on d-DNNF.
Given a literal ranking function $r(\ell)$ that assign ranks
to each literal $\ell$, one defines the rank $r(\omega)$ of a
model $\omega$ as the sum of the ranks of the literals made true
by $\omega$ (i.e., $r(\omega)\doteq\sum_{\omega\vDash\ell}r(\ell)$),
and say that $\omega$ is a best (ranked) model if there is no model
$\omega'$ such that $r(\omega')<r(\omega)$.

Given a theory in d-DNNF, one computes the rank of the best models,
and a best model, in time linear in the size of the d-DNNF.
This computation transforms the DAG of the d-DNNF into a arithmetic
circuit by replacing the AND nodes with `+' and the OR nodes with `min'.
The literal ranking function assigns
values to the leaves of the circuit that are propagated to the
root in linear time. The value of the root is the rank of the
best model \cite{darwiche:weighted}.

In this paper, we exploit the properties of d-DNNFs by constructing
a logical theory whose models encode the valid workflow instantiations
and best models encode the optimal (best) instantiations.
Thus, the combinatorial search is reduced to the computation
of a best model of a logic theory which can be efficiently
performed once the theory is transformed into d-DNNF format.

The paper contains six more sections. The next section illustrates and
motivates the WIP with a simple yet typical example. Section~3 summarizes
the existing related work in three areas of selection of services,
query rewriting, and knowledge compilation.
Then, Sections~4 and 5 describe the architecture of the system and 
report our empirical results over different benchmark problems
respectively.
For the interested reader, Section~6 presents a formal description
of the proposed approach together with an analysis.
The paper concludes with a final discussion in Section~7.

\section{Motivating Example}

Consider a simple flight-information system which contains information
about flights between cities and information about which cities are in
the US. Such a system can be described using LAV with the two abstract
services $\flight(x,y)$ and $\UScity(x)$. The former relates two cities
$x$ and $y$ if there is a direct flight between them, and the latter tells
whether $x$ is a US city or not.

For the concrete services, assume that the available data sources on
the Internet contain the following information:
\begin{enumerate}[--]
\item $\national(x,y)$ relates two US cities that are connected by a direct flight,
\item $\oneway(x,y)$ relates two cities that are connected by a one-way flight,
\item $\onestop(x,y)$ relates two cities that are connected by a one-stop flight,
\item $\flightPA(x)$ tells if there is a direct flight from $x$ to Paris,
\item $\onestopPA(x,y)$ relates $x$ and $y$ if there is a flight from $x$ to Paris
      with a stop at $y$, and
\item $\fromNY(x)$ tells if there is a flight from New York into $x$.
\end{enumerate}
Furthermore, the concrete services are described using the abstract services:
\begin{alignat*}{1}
\national(x,y)\   &\qrule\ \flight(x,y),\,\UScity(x),\,\UScity(y)\,. \\
\oneway(x,y)\     &\qrule\ \flight(x,y)\,. \\
\onestop(x,z)\    &\qrule\ \flight(x,y),\,\flight(y,z)\,. \\
\flightPA(x)\     &\qrule\ \flight(x,\PA)\,. \\
\onestopPA(x,y)\  &\qrule\ \flight(x,y),\,\flight(y,\PA)\,. \\
\fromNY(x)\       &\qrule\ \flight(\NY,x)\,.
\end{alignat*}
Now suppose that a user is interested in constructing a workflow able to retrieve
the one-stop round-trip flights from US cities to any city in the world, such that
flights can stop at any city. 
\[ W(x,w,y,z) \qrule \UScity(x),\,\flight(x,w),\,\flight(w,y),\,\flight(y,z),\,\flight(z,x)\,. \]
The following conjunctive query represents the workflow that defines this request
in terms of abstract services. The workflow is defined in way that all issues
about the binding of input/output parameters between services had been resolved
such that any instantiation of the abstract services in terms of concrete services
is a valid implementation of the workflow. 
Implementations correspond to compositions of concrete services in which a concrete
service may implement one or more abstract services from the workflow, but each
abstract service can be implemented by exactly one concrete service.
For example, the following composition corresponds to one such implementation.
\[ I(x,w,y,z)\ \qrule\ \national(x,w),\,\flightPA(w),\,\oneway(\PA,z),\,\national(z,x)\,. \]
However, the following two compositions are not valid.
\begin{alignat*}{1}
I'(x,w,y,z)\  \qrule\ &\national(x,y),\,\flightPA(y),\,\fromNY(z),\,\national(z,x)\,. \\
I''(x,w,y,z)\ \qrule\ &\onestop(x,y),\,\oneway(y,z),\,\national(z,x)\,.
\end{alignat*}
The first composition is not valid because it maps the workflow variable $y$
into constants \PA\ and \NY\ that denote different cities.
On the other hand, $I''$ does not implement the workflow because the concrete
service $\onestop(x,y)$ does not receive as input, or produce as output, the
middle city where the flight stops and thus is not possible to ensure that 
this city is bound to the city $w$ that is returned by the workflow.

These examples show that a proper workflow instantiations must handle
constants such that not two different constants are mapped to each
other either directly or indirectly via transitivity, and that all
attributes that appear in a join or in the output need to be produced
by the selected concrete services.

The QoS parameters are modeled by annotating the concrete services with
utilities that characterize their behavior which are then aggregated
during instantiation. Thus, as said before, the best instantiation is
the one that minimizes (or maximizes) the aggregation of utilities.

\section{Related Work}

In this section we summarize the existing approaches that provide
solutions to the problems of service selection, query rewriting and
discuss related work in the area of Artificial Intelligence called
knowledge compilation.

\subsection{Service Selection}

The problem of selecting the services that implement an abstract workflow
and best fit the QoS-based criteria is known as the QoS-aware service
selection or composition problem, which has been shown to be NP-hard \cite{Hiroshi2008}.
This problem is a combinatorial optimization problem and several heuristics 
have been proposed to find a relatively good solution in a reasonably short
period of time.

Rahmani et al.\ \cite{rahmani08} present a distance metric-based heuristic
that focus a backward search algorithm; this metric induces an order of the
services in a way that sink nodes are unlikely to be visited.
In a series of papers, Berardi and others \cite{berardi05,berardi08,berardi06}
describe services and workflows in terms of deterministic finite-state machines
that are encoded using Description Logic theories whose models correspond to
solutions of the problem. Although reasoning methods for Description Logics
formalisms could be exploited, scalability or performance of the proposed
solution has not been reported.

Ko et al.\ \cite{myoung08} propose a constraint-based approach that encodes the
non-functional permissible values as a set of constraints whose violation needs
to be minimized; to traverse the space of possibly optimal solutions, a hybrid
algorithm that combines tabu search and simulated annealing meta-heuristics is
implemented. Experimental results show that the proposed solution is able to
scale up to a large number of services and abstract processes.
Cardellini et al.\ \cite{cardellini07} encode one part of the QoS-aware service
composition problem as a Linear Programming problem \cite{cardellini07}.
On the other hand, Wada el at.\ \cite{Hiroshi2008} treat the problem as a 
multi-objective optimization problem where the different QoS parameters are
considered equally important instead of aggregating them into a single function.
Then, a genetic-based algorithm is proposed to identify a set of non-dominated
service compositions that best fit all the QoS requirements.

Alrifai and Risse \cite{alrifaiR09} propose a two-fold solution that uses a
hybrid integer programming algorithm to find the decomposition of global QoS
into local constraints, and then, selects the services that best meet the local
constraints.

Recently, two planning-based approaches have been proposed.
Kuter and Golbeck \cite{kuterG09} extend the SHOP2 planning algorithm to select
the trustworthy composition of services that implement a given OWL-S process model,
while Sohrabi and McIlraith \cite{sohrabiM09} propose a HTN planning-based solution
where user preference metrics and domain regulations are used to guide the planner
into the space of relevant compositions.
Finally, L\'ecu\'e \cite{lecue09} proposes a genetic-based algorithm to identify
the composition of services that best meet the quality criteria for a set of QoS
parameters.

Although these solutions are able to solve the optimization problem and scale up
to a number of abstract processes, none of them are tailored to semantically describe
the services in terms of abstract process, nor to use these descriptions to 
identify the services that implement a given workflow or best meet the user's
non-functional criteria.

\subsection{Query Rewriting}

A number of algorithms have been developed to find the rewritings of a given query;
the most prominent being the bucket algorithm \cite{levy:bucket}, the inverse rules
algorithm \cite{pods:DuschkaG97,Qian96}, the \minicon\ algorithm \cite{pottinger:minicon},
and the \mcdsat\ algorithm \cite{arvelo:aaai06}.
Generally, query rewriting algorithms work in two phases. During the first phase,
the algorithms identify the views that rewrite at least one sub-goal of the query, and
during the second, these partial rewritings are combined into complete rewritings.
The main difference between the algorithms is the criteria used to choose the relevant
views to reduce the space of non-useful rewritings.

The bucket algorithm reduces the number of possibilities by just considering each
sub-goal in the query in isolation, and selecting the views that are able to produce
at least the attributes projected by the query. Since the attributes involved in the
joins in the query are not verified, a large number of rewritings comprised of Cartesian
products may be generated.

The Inverse Rules algorithm constructs a set of rules that invert the view definitions
and establish how to compute tuples for the database relations from the tuples of the
views. Similarly to the bucket algorithm, it can produce a large number of non-useful
rewritings.

The \minicon\ algorithm overcomes the limitations of the previous algorithms by identifying
only views that rewrite a set of the sub-goals of the query, and that can be combined with
the rest of the sub-goals. The key idea is to identify the mappings between the variables in
each sub-goal to the variables in one or more sub-goals in the views, in a way
that, join variables in the query are mapped to join variables in the body of a view or to
the distinguished variables of the view. Mappings between variables and sub-goals are
represented in the so-called MiniCon Descriptions (MCDs) \cite{pottinger:minicon}.

Finally, the \mcdsat\ algorithm is able to identify the query rewritings of a query by
translating the problem of rewriting into the problem of enumerating the models of a
propositional theory whose models are in correspondence with the rewritings of the query.
The algorithm exploits the properties of d-DNNFs to efficiently enumerate the models of
the theory.
The \mcdsat\ algorithm has demonstrated to scale better than the \minicon\ algorithm over
a large number of benchmarks often showing performance improvements of several orders of
magnitude. However, the \mcdsat\ algorithm was not designed for rewriting problems
involving explicit constants, nor to compute the best rewritings with respect to a given
utility function or cost model. In this paper, we propose an extended encoding that 
overcomes these limitations and apply the encoding to the Workflow Instantiation Problem.

\subsection{Knowledge Compilation}

Knowledge compilation is the area in AI concerned with the problem of mapping
logical theories into suitable fragments that make certain desired operations
tractable \cite{cadoli:compilation}. Different compilation languages have been
defined, for instance, Ordered Binary Decision Diagrams (OBDDs) \cite{bryant:obdd},
Negation Normal Form (NNF) \cite{barwise:handbook}, and Decomposable Negation
Normal Form (DNNF) \cite{darwiche:map}. In this work, we make use of the properties
of the deterministic DNNFs (d-DNNF) \cite{darwiche:d-dnnfs} to provide an scalable
and efficient solution to the service selection problem.

\begin{figure}[t]
\centering
\includegraphics[width=.7\textwidth]{odd}
\caption{A decomposable and deterministic NNF.}
\label{fig:dnnf}
\end{figure}
